Capstone Movielens Report

### 2025-12-06 reports/capstone-movielens.site/05.appendix.Rmd
Function described: calc_UMGY_effect_MSE.

The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-2]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_UMGY_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.

@@

#### [calc_UMGY_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L112) Function {#func.calc_UMGY_effect_MSE}
\

Validates the *UMGY Effect (UMGYE) Model* produced by the  [train_user_movie_genre_effect](#func.train_user_movie_genre_effect) or [train_user_movie_genre_effect.cv](#func.train_user_movie_genre_effect.cv) function call.

@@

@@

:::
::: { #func.calc_UMGY_effect_MSE.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMGY Effect Model*.

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGYE Model* passed in the [umg_effect](#func.calc_UMGY_effect_MSE.args.umg_effect) argument using the validation dataset passed in the [test_set](#func.calc_UMGY_effect_MSE.args.test_set) arg.

@@

The source code of the [calc_UMGY_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L112) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [calc_UMGY_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L112) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L112) script on _GitHub_.
:::
:::

@@

::: { #func.calc_UMGY_effect_MSE.cv.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMGY Effect Model*.

@@

The function calculates *MSE* for the *UMGYE Model* passed in the [umg_effect](#func.calc_UMGY_effect_MSE.cv.args.umg_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

::: { #func.calc_UMGY_effect_RMSE.cv.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMGY Effect Model*.

@@

The function calculates *Root Squared Error (RMSE)* for the *UMGYE Model* passed in the [umg_effect](#func.calc_UMGY_effect_RMSE.cv.args.umg_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

The simplified version of the source code of the [regularize.test_lambda.UMGY_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L148) function is shown below:
```{r eval=FALSE}
regularize.test_lambda.UMGY_effect.cv <- function(lambda){
  umg_effect <- train_user_movie_genre_effect.cv(lambda)
  calc_user_movie_genre_effect_RMSE.cv(umg_effect)
}
```

>>>>

Under the hood, this function internally calls the [calc_date_general_effect](#func.calc_date_general_effect) and [calc_UMGY_effect](#func.calc_UMGY_effect) functions (both described above in *This Section*).

::: {.noteblock data-latex=""}
The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-2]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_UMGY_effect.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

#### [calc_UMGY_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L94) Function {#func.calc_UMGY_effect_MSE}
\

Validates the *UMGY Effect (UMGYE) Model* produced by the  [train_UMGY_effect](#func.train_UMGY_effect) or [train_UMGY_effect.cv](#func.train_UMGY_effect.cv) function call.

::: {.noteblock data-latex=""}
This helper function is used internally in the [calc_UMGY_effect_MSE.cv](#func.calc_UMGY_effect_MSE.cv) function (described below in *This Section*) as shown in the [Usage](#func.calc_UMGY_effect_MSE.usage) subsection of *This Function's Description Section* below.
:::

@@

```{r eval=FALSE}
    cv_fold_dat$validation_set |> calc_UMGY_effect_MSE(UMGY_effect)
```

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_UMGY_effect_MSE` is called internally from the [calc_UMGY_effect_MSE.cv](#func.calc_UMGY_effect_MSE.cv) function described below in *This Section*.
:::
:::

@@

:::
::: { #func.calc_UMGY_effect_MSE.args.UMGY_effect .sidebar }
  - **UMGY_effect:** Data frame representing the *UMGY Effect Model*.

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGYE Model* passed in the [UMGY_effect](#func.calc_UMGY_effect_MSE.args.UMGY_effect) argument using the validation dataset passed in the [test_set](#func.calc_UMGY_effect_MSE.args.test_set) arg. 
:::

@@

The source code of the [calc_UMGY_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L94) function is shown below:
```{r eval=FALSE}
calc_UMGY_effect_MSE <- function(test_set, UMGY_effect){
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(UMGY_effect, by='year') |>
    mutate(resid = rating - clamp(mu + a + b + g + ye)) |> 
    pull(resid) |> mse()
}
```

::: {.noteblock data-latex=""}
The source code of the [calc_UMGY_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L94) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L94) script on _GitHub_.
:::
:::

@@

::: { #func.calc_UMGY_effect_MSE.cv.args.UMGY_effect .sidebar }
  - **UMGY_effect:** Data frame representing the *UMGY Effect Model*.

@@

The function calculates *MSE* for the *UMGYE Model* passed in the [UMGY_effect](#func.calc_UMGY_effect_MSE.cv.args.UMGY_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

::: { #func.calc_UMGY_effect_RMSE.cv.args.UMGY_effect .sidebar }
  - **UMGY_effect:** Data frame representing the *UMGY Effect Model*.

@@

The function calculates *Root Squared Error (RMSE)* for the *UMGYE Model* passed in the [UMGY_effect](#func.calc_UMGY_effect_RMSE.cv.args.UMGY_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

The simplified version of the source code of the [regularize.test_lambda.UMGY_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L148) function is shown below:
```{r eval=FALSE}
regularize.test_lambda.UMGY_effect.cv <- function(lambda){
  umgy_effect <- train_user_movie_genre_effect.cv(lambda)
  calc_user_movie_genre_effect_RMSE.cv(umgy_effect)
}
```
